---
title: "Assignment 2 - ACM-S22"
output: html_notebook
author: ""
---

# Description
For this assignment, we implement an imperfect Win-Stay Lose-Shift (WSLS) model,
simulate data and fit the model in Stan. Furthermore, parameter recovery is
assessed and discussed.

### Instructions:
1. Simulate appropriate data (1 agent, 1 opponent, n trials)
2. Code the model in Stan
3. Fit the model
4. Simulate a full range of data (n agents, t trials,
possibly different opponents)
5. Fit the model separately to each agent
6. Assess parameter recovery
7. Assess the importance of the amount of trials: does parameter recovery
change with 10, 30, 50, 100, 1000 trials?  

**Optional:** build the same model in brms, extract the Stan code from it
using `(stancode(model))` and comment the lines of code that you recognize
and understand.



# Setup
We rely on Stan using the `cmdstanr` R package.
Furthermore, we import some utility functions such as agent definitions and functions for running simulations.

The `utils/simulation.R` contains the `simulation()` function, which runs a single simulation and returns two vectors:

* `random_face`, which represents the choices (heads/tails) of the random agent  
* `WSLS_face`, which represents the choices (heads/tails) of the WSLS agent  

The `utils/simulation_helper_functions.R` file contains the `WSLS_encoding()` function. This function encodes wins/losses and heads/tails into two new vectors (for use in the Stan model). The format for the function is:

* `wins_face`: if (win & heads) then 1, if (win & tails) then -1, else 0  
* `loss_face`: if (loss & heads) then -1, if (loss & tails) then 1, else 0  

```{r Setup, message=FALSE}

library(pacman)

pacman::p_load(tidyverse,
               here,
               posterior,
               cmdstanr,
               brms,
               ggridges)

# Define paths for folders
plots_path = "plots"
stan_path  = "stan_models"
utils_path = "utils"

# Source external script for agent definitions
# 1. The imperfect WSLS agent (main focus of analysis)
# 2. A random-bias agent, representing the opponent
source("utils/agents.R")
source("utils/simulation.R")
source("utils/simulation_helper_functions.R")
source("utils/create_posterior_draws.R")

```



# Run the simulation with a single agent
The Stan model is located at `stan_models/WSLS.stan`.

```{r Single agent simulation}

# Define parameters for the simulation
trials          <- 120
random_bias     <- .5
WSLS_staybias   <- 0.8
WSLS_leavebias  <- 0.5

# Running the simulation
single_agent_df <- simulation_to_df(
  simulation(
    trials, random_bias, WSLS_staybias, WSLS_leavebias))



# -- Stan modeling --
file <- file.path(paste(stan_path, "\\WSLS.stan", sep=""))
single_agent_stan <- cmdstan_model(file, cpp_options = list(stan__threads = TRUE))

data <- list(
  trials    = trials,
  face      = single_agent_df$WSLS_face,
  loss_face = single_agent_df$loss_face,
  wins_face = single_agent_df$wins_face)

# The following command calls Stan with specific options.
single_agent_samples <- single_agent_stan$sample(
  data = data,
  seed = 123,
  chains = 2,
  parallel_chains = 2,
  threads_per_chain = 2,
  iter_warmup = 1000,
  iter_sampling = 2000,
  refresh = 500,
  max_treedepth = 20,
  adapt_delta = 0.99,
)

single_agent_samples$summary()

```



# Run the simulation with multiple agents
The Stan model is located at `stan_models/WSLS_multiagent.stan`.

```{r Multi agent simulation}

# Define parameters for the simulation
trials          <- 100
agents          <- 10
random_bias     <- .55
WSLS_staybias   <- 0.9
WSLS_leavebias  <- 0.8

# Prepare the data for multiple agents
for (agent in seq(agents)) {
  # Running a single simulation
  single_sim_result <- simulation_to_df(
    simulation(
      trials, random_bias, WSLS_staybias, WSLS_leavebias), agent)
  
  # Conditional for creating 
  if (agent == 1) {
    multi_agent_df <- single_sim_result
  } else {
    multi_agent_df <- rbind(multi_agent_df, single_sim_result)
  }
  
  # Cleanup environment
  rm(single_sim_result, agent)
}



# -- Stan modeling --
file <- file.path(paste(stan_path, "\\WSLS_multiagent.stan", sep=""))
multi_agent_stan <- cmdstan_model(file, cpp_options = list(stan__threads = TRUE))

data <- list(
  trials    = trials,
  agents    = agents,
  face      = matrix(multi_agent_df$WSLS_face, nrow=trials, ncol=agents),
  wins_face = matrix(multi_agent_df$wins_face, nrow=trials, ncol=agents),
  loss_face = matrix(multi_agent_df$loss_face, nrow=trials, ncol=agents))

# The following command calls Stan with specific options.
multi_agent_samples <- multi_agent_stan$sample(
  data = data,
  seed = 123,
  chains = 2,
  parallel_chains = 2,
  threads_per_chain = 2,
  iter_warmup = 1000,
  iter_sampling = 2000,
  refresh = 500,
  max_treedepth = 20,
  adapt_delta = 0.99,
)

multi_agent_samples$summary()

```



# Assess parameter recovery
We want to asses parameter recovery, that is, the ability of our model estimates to approach the actual parameters. Since we have defined the parameters ourselves, we can simulate varying number of trials and the influence on our estimates for the parameters.

```{r Parameter recovery}

multi_agent_draws <- create_draws_df(multi_agent_samples)

# Plot a ridgeline plot for staybias probability estimates
multi_agent_draws %>%
  subset(bias == "staybias") %>%

  ggplot(aes(x=posterior_estimate, y=agent, fill=..y.., color="white")) +
  geom_density_ridges(rel_min_height = 0.01) +
  scale_fill_gradient(low="deepskyblue4", high="deepskyblue3") +
  scale_color_manual(values = c("white")) + 
  geom_vline(xintercept=WSLS_staybias, size=1.1, color="firebrick3", alpha=0.7) +
  xlim(0, 1) +
  
  theme_ridges() + 
  theme(legend.position = "none") +
  labs(
    title="Staybias posterior probability estimates",
    x="Staybias posterior probability",
    y="")

ggsave(paste(plots_path, "\\multi_agent_staybias.jpg", sep=""), width=7.29, height=4.5)


# Plot a ridgeline plot for leavebias probability estimates
multi_agent_draws %>%
  subset(bias == "leavebias") %>%
  
  ggplot(aes(x=posterior_estimate, y=agent, fill=..y.., color="white")) +
  geom_density_ridges(rel_min_height = 0.01) +
  scale_fill_gradient(low="seagreen4", high="seagreen3") +
  scale_color_manual(values = c("white")) + 
  geom_vline(xintercept=WSLS_leavebias, size=1.1, color="firebrick3", alpha=0.7) +
  xlim(0, 1) +
  
  theme_ridges() + 
  theme(legend.position = "none") +
  labs(
    title="Leavebias posterior probability estimates",
    x="Leavebias posterior probability",
    y="")

ggsave(paste(plots_path, "\\multi_agent_leavebias.jpg", sep=""), width=7.29, height=4.5)

```



```{r Parameter recovery across different trials}

# Define parameters for the simulation
trials_vector   <- c(10, 30, 50, 100, 1000)
agents          <- 10
random_bias     <- .55
WSLS_staybias   <- 0.9
WSLS_leavebias  <- 0.8

# Stan model setup
file <- file.path(paste(stan_path, "\\WSLS_multiagent.stan", sep=""))
multi_agent_stan <- cmdstan_model(file, cpp_options = list(stan__threads = TRUE))

# Go through each number of trials and save posterior draws
for (trials in trials_vector) {
  # Prepare the data for multiple agents
  for (agent in seq(agents)) {
    # Running a single simulation
    single_sim_result <- simulation_to_df(
      simulation(
        trials, random_bias, WSLS_staybias, WSLS_leavebias), agent)
    
    # Conditional for creating 
    if (agent == 1) {
      multi_agent_df <- single_sim_result
    } else {
      multi_agent_df <- rbind(multi_agent_df, single_sim_result)
    }
    
    # Cleanup environment
    rm(single_sim_result, agent)
  }
  
  data <- list(
    trials    = trials,
    agents    = agents,
    face      = matrix(multi_agent_df$WSLS_face, nrow=trials, ncol=agents),
    wins_face = matrix(multi_agent_df$wins_face, nrow=trials, ncol=agents),
    loss_face = matrix(multi_agent_df$loss_face, nrow=trials, ncol=agents))
  
  # The following command calls Stan with specific options.
  multi_agent_samples <- multi_agent_stan$sample(
    data = data,
    seed = 123,
    chains = 2,
    parallel_chains = 2,
    threads_per_chain = 2,
    iter_warmup = 1000,
    iter_sampling = 2000,
    refresh = 500,
    max_treedepth = 20,
    adapt_delta = 0.99,
  )
  
  # Get all the posterior samples, save to dataframe and convert to long format
  multi_agent_draws <- create_draws_df(multi_agent_samples)
  multi_agent_draws$trials <- trials
  
  # Combine in a dataframe
  if (trials == trials_vector[1]) {
    multi_trial_draws <- multi_agent_draws
  } else {
    multi_trial_draws <- rbind(multi_trial_draws, multi_agent_draws)
  }
  
}

# Setup factor ordering for trials
multi_trial_draws$trials <- as.factor(multi_trial_draws$trials)
levels(multi_trial_draws$trials) <- trials_vector # Or reverse: rev.default(trials_vector)


multi_trial_draws %>%
  subset(bias == "staybias") %>%
  
  ggplot(aes(x=posterior_estimate, y=trials, fill=..y.., color="white")) +
  geom_density_ridges(rel_min_height = 0.01, scale=2.5) +
  scale_fill_gradient(low="deepskyblue4", high="deepskyblue3") +
  scale_color_manual(values = c("white")) + 
  geom_vline(xintercept=WSLS_staybias, size=1.1, color="firebrick3", alpha=0.7) +
  xlim(0, 1) +
  
  theme_ridges() + 
  theme(legend.position = "none") +
  labs(
    title="Staybias posterior probability estimates by number of trials",
    x="Staybias posterior probability",
    y="Number of Trials")

ggsave(paste(plots_path, "\\multi_trial_staybias.jpg", sep=""), width=7.29, height=4.5)


# Plot a ridgeline plot for leavebias probability estimates
multi_trial_draws %>%
  subset(bias == "leavebias") %>%
  
  ggplot(aes(x=posterior_estimate, y=trials, fill=..y.., color="white")) +
  geom_density_ridges(rel_min_height = 0.01, scale=2.5) +
  scale_fill_gradient(low="seagreen4", high="seagreen3") +
  scale_color_manual(values = c("white")) + 
  geom_vline(xintercept=WSLS_leavebias, size=1.1, color="firebrick3", alpha=0.7) +
  xlim(0, 1) +
  
  theme_ridges() + 
  theme(legend.position = "none") +
  labs(
    title="Leavebias posterior probability estimates by number of trials",
    x="Leavebias posterior probability",
    y="Number of Trials")

ggsave(paste(plots_path, "\\multi_trial_leavebias.jpg", sep=""), width=7.29, height=4.5)

```





